Development of a 3D display

### 2016-05-16 paper.tex
Unindented \itemize, removed trailing \s
Another chapter, nicer math, etc

\usepackage{xspace}
\def \iZc {I\textsuperscript{2}C\xspace}
\def \vdd {V\textsubscript{DD}}

@@

% Display SI units
\usepackage{siunitx}

@@

As far as I know, however, only two people have built a working version of a three-dimensional
POV display so far (and have published it on the internet).\footnote{It turned out that there was
a third one: When I posted a picture of my project on the internet, a reddit user told me about
the display he built years before.}
Neither of these is being produced on a grand scale, instead they are --- like most 2D versions
--- side projects of makers or engineers.

@@

	\item Resolution of 16 * 100 * 10 pixels (radius * circumference * height)
	\item Each pixel consists of an RGB LED
	\item Colour depth of 3 bit per pixel, or one bit per LED
	\item Rotational frequency of at least 30 Hz, or as fast as it needs to be in order to not flicker

@@

	\item Controlled using \iZc  instead of SPI

Because \iZc uses pullup resistors instead of a push-pull drive, it takes a little moment for the
logic level to return to its normal state. In order to ensure correct data transmission, it is
necessary to set a relatively slow clock speed. The two \iZc interfaces of the Teensy only support
clock speeds of 2.4 MHz --- and because \iZc requires some overhead data to send out the slave's
addresses, the actual data rate will be even lower.

@@

\begin{ownmath}
% Yes it's a mess but \si makes it a unit, so it is not in cursive while still being small enough
24\,\frac{\si{bit}}{\si{pixel}} * 16000\,\frac{\si{pixels}}{\si{rotation}} * 30\,\frac{\si{rotations}}{\si{second}} = \SI{11.52}{Mbps}
\end{ownmath}

This means that I'd need at least 5 parallel \iZc lines to control all the LEDs with 8 bit colour
depth. Because each Teensy only has two of them, I'd need three Teensy boards, which would be
possible but very cumbersome.

@@

After quite some searching around, I decided to use the TLC5927 by Texas Instruments. This is
essentially a shift register with some extra features. Like every shift register, it has a serial
data input (SDI), a serial data output (SDO), and a clock pin (CLK). Additionally, there is a latch
pin (LE), which serves the purpose of only refreshing the state of the LEDs once all data has been
sent.

@@

	\item Colours: Red, green and blue
	\item Current: 20 mA or more. Because the LEDs are within one pixel only $\frac{1}{100}$ of the
time, they are essentially pulse width modulated, reducing their effective brightness. In order to
still see the image clearly without darkening the room, they have to be bright enough.
	\item Mounting type: Surface mount, so that I can route traces past the LEDs on the other side
of the PCB
	\item Size: At most 5 * 5 mm, preferably smaller, but still large enough to be able to solder them
by hand
	\item Diffused lens, so the colours mix well even when viewing the LED directly instead of using
it for illumination.

@@

These requirements reduced the category "LED Indication - Discrete" from 18940 to 130 products.
From those remaining products, I chose the one that was cheapest in a quantity of 160. This
led me to an LED with the beautiful name CLVBA-FKA-CAEDH8BBB7A363.

@@

% Lots of ugly hacks to have a non-italic font
\begin{ownmath}
\textrm{\iout} = 15 * \frac{\SI{1.25}{\volt}}{\si{\rext}}
\end{ownmath}

@@

\begin{ownmath}
\textrm{\rext} = 15 * \frac{\SI{1.25}{\volt}}{\SI{20}{\milli\ampere}} = \SI{937.5}{\ohm}
\end{ownmath}

@@

In order to achieve this, but also to avoid large gaps between the pixels, I placed the LEDs as
close as possible to each other. The datasheet claims a width of \SI{2.8 +- 0.2}{\milli\meter},
so I decided to place the LEDs at \SI{3}{\milli\metre} increments. On a grid of \SI{12.5}{mil},
the next larger distance is \SI{125}{mil}, which is equal to \SI{3.175}{mm}.

@@

\begin{ownmath}
\textrm{P\textsubscript{MAX}} = (\SI{5}{\volt} - \textrm{V\textsubscript{F}}) *
\SI{20.6}{\milli\ampere\per output} * \SI{16}{outputs} = \SI{780}{\milli\watt}
\end{ownmath}

@@

To save the image on the microcontroller, I used a simple three-dimensional array containing 100 *
10 * 6 bytes of data. The first index stands for the angle of a pixel in hundredths of a circle,
the second index corresponds to the height and the third one describes the position of each byte 
on a row of LEDs. Because there are 48 LEDs per row, 6 bytes can be used to describe the row's
state without wasting any storage space. The only way to be more efficient in terms of storage
would be to compresss data, which would require significantly more computing power.

@@

>>>>

\usepackage{xspace}
\def \itwoc {I\textsuperscript{2}C\xspace}
\def \vdd {V\textsubscript{DD}}

@@

% Display SI units
\usepackage[separate-uncertainty=true, multi-part-units=single]{siunitx}

@@

As far as I know, however, only two people have built a working version of a three-dimensional
POV display so far (and have published it on the internet).\footnote{It turned out that there was
a third one: When I posted a picture of my project on the internet, a reddit user told me about
the display he built years before.}
Neither of these is being produced on a grand scale, instead they are --- like most 2D versions
--- side projects of hobbyists or engineers.

@@

\item Resolution of $16 \cdot 100 \cdot 10$ pixels (radius $\cdot$ circumference $\cdot$ height)
\item Each pixel consists of an RGB LED
\item Colour depth of 3 bit per pixel, or one bit per LED
\item Rotational frequency of at least 30 Hz, or as fast as it needs to be in order to not flicker

@@

\item Controlled using \itwoc  instead of SPI

Because \itwoc uses pullup resistors instead of a push-pull drive, it takes a little moment for
the logic level to return to its normal state. In order to ensure correct data transmission, it is
necessary to set a relatively slow clock speed. The two \itwoc interfaces of the Teensy only
support clock speeds of up to 2.4 MHz --- and because \itwoc requires some overhead data to send
out the slave's addresses, the actual data rate will be even lower.

@@

\begin{ownmath}
% Yes it's a mess but \si makes it a unit, so it is not in cursive while still being small enough
24\,\frac{\si{bit}}{\si{pixel}} \cdot 16000\,\frac{\si{pixels}}{\si{rotation}} \cdot
30\,\frac{\si{rotations}}{\si{second}} = \SI{11.52}{Mbps} 
\end{ownmath}

This means that I'd need at least 5 parallel \itwoc lines to control all the LEDs with 8 bit colour
depth. Because each Teensy only has two of them, I'd need three Teensy boards, which would be
possible but very cumbersome.

@@

After quite some searching around, I decided to use the TLC5927 by Texas Instruments. It is
essentially a shift register with some extra features. Like every normal shift register, it has a
serial data input (SDI), a serial data output (SDO), and a clock pin (CLK). Additionally, there is
a latch pin (LE), which serves the purpose of only refreshing the state of the LEDs once all data
has been sent.

@@

\item Colours: Red, green and blue

\item Current: 20 mA or more. Because the LEDs are within one pixel only $\frac{1}{100}$ of the
time, they are essentially pulse width modulated, reducing their effective brightness. In order to
still see the image clearly without darkening the room, they have to be bright enough.

\item Mounting type: Surface mount, so I can still use the other side of the board

\item Size: At most $\SI{5}{\milli\meter}$ by $\SI{5}{\milli\meter}$, preferably smaller, but
still large enough to solder by hand

\item Diffused lens, so the colours mix well even when viewing the LED directly instead of using
it for illumination.

@@

These requirements reduced the category "LED Indication - Discrete" from 18940 to 130 items.
From those remaining products, I chose the one that was cheapest in a quantity of 160. This
led me to an LED with the beautiful name CLVBA-FKA-CAEDH8BBB7A363.

@@

% Lots of ugly hacks to have a non-italic font
\begin{ownmath}
\textrm{\iout} = 15 \cdot \frac{\SI{1.25}{\volt}}{\si{\rext}}
\end{ownmath}

@@

\begin{ownmath}
\textrm{\rext} = 15 \cdot \frac{\SI{1.25}{\volt}}{\SI{20}{\milli\ampere}} = \SI{937.5}{\ohm}
\end{ownmath}

@@

In order to achieve this, but also to avoid large gaps between the pixels, I placed the LEDs as
close as possible to each other. The datasheet claims a width of \SI{2.8 \pm 0.2}{\milli\meter},
so I decided to place the LEDs at \SI{3}{\milli\metre} increments. On a grid of \SI{12.5}{mil},
the next larger distance is \SI{125}{mil}, which is equal to \SI{3.175}{mm}.

@@

\begin{ownmath}
\textrm{P\textsubscript{MAX}} = (\SI{5}{\volt} - \textrm{V\textsubscript{F}}) \cdot
\SI{20.6}{\milli\ampere\per output} \cdot \SI{16}{outputs} = \SI{780}{\milli\watt}
\end{ownmath}

@@

To save the image on the microcontroller, I used a simple three-dimensional array containing $100
\cdot 10 \cdot 6$ bytes of data. The first index stands for the angle of a pixel in hundredths of
a circle, the second index corresponds to the height and the third one describes the position of
each byte on a row of LEDs. Because there are 48 LEDs per row, 6 bytes can be used to describe the
row's state without wasting any storage space. The only way to be more efficient in terms of
storage would be to compresss data, which would require significantly more computing power.

@@

\chapter{Operation}

\section{Power Supply}

Initially I had planned to supply the PCBs with \SI{5}{\volt}. However, the Teensy sends its
signals at \SI{3.3}{\volt}. The driver chips, on the other hand, only read a logical high when a
signal is over $0.7 \cdot \textrm{\vdd}$. At \SI{5}{\volt}, this value would be \SI{3.5}{\volt}.
So I tried powering the boards with only \SI{3.3}{\volt}, and everything worked perfectly. Better
yet, the boards weren't getting as hot anymore.

According to the manufacturer, the Teensy needs at least \SI{3.7}{\volt} to operate, presumably
because it uses a linear regulator to get a stable voltage ov \SI{3.3}{\volt}. However, there were
no problems when I started using \SI{3.3}{\volt}. Even though the regulator probably reduces that
voltage to less than \SI{3}{\volt}, the processor still runs at its full clock speed of
\SI{96}{\mega\hertz}.

Unfortunately, the PC power supply which I used for the project broke down. As I later found out,
the culprit was the motor consuming too much current. At that time, I didn't use any series
resistance to limit the current flowing through the motor, which I powered with \SI{3.3}{\volt}.
This has caused the inrush current to be greater than what the power supply could handle.

To remedy this, I used a pair of long, thin wires to connect to the motor to provide some
resistance. I later replaced these with two \SI{150}{\milli\ohm} power resistors. Now there are two
ways to power the motor: I can use \SI{5}{\volt} and connect the resistors in series for a total
resistance of \SI{300}{\milli\ohm}, or I can power it with \SI{3.3}{\volt} and connect the
resistors in parallel, resulting in a resistance of \SI{75}{\milli\ohm}. The latter method is more
efficient, but it causes the acceleration at startup to be slower.

\section{Mechanical Problems}

Because the slip rings weren't fixed in their position, they could slide out of their slots after
a few minutes of operation. I fixed this issue by mounting a small wooden board onto these slots
(perventing lateral movement), and routing the cables going from the power supply to the sliprings
through two holes in that board (perventing lengthwise movement).

The boards weren't exactly aligned at the \SI{28.8}{\degree} that I specified. This caused the
image to be displayed crookedly: Pixels that should've been right above each other were shifted
sideways by a small distance. Whenever I adjusted the boards, they would be misaligned again after
using the display for some time. To fix this, I aligned them correctly one last time and used
enough solder on the power cables running through the boards for them to form a solid connection.

The bolt nut that held the motor in place started coming loose because of the vibrations. To
pervent that from happening again, I used a second nut and tightened it against the first one.
